\chapter{List of publications}

\section*{International journal papers (published)}

\begin{itemize}
\item Claesen, M., De Smet, F., Suykens, J. A. K., \& De Moor, B. (2014). ``\emph{EnsembleSVM: A library for ensemble learning using support vector machines}''. Journal of Machine Learning Research, 15(1), 141--145.
\item Claesen, M., De Smet, F., Suykens, J. A., \& De Moor, B. (2015). ``\emph{A robust ensemble approach to learn from positive and unlabeled data using SVM base models}''.
Neurocomputing, 160, 73-84.
\end{itemize}

\section*{International journal papers (submitted)}
\begin{itemize}
\item Claesen, M., De Smet, F., Gillard, P., Mathieu, C.\& De Moor, B. (2015). ``\emph{Building Classifiers to Predict the Start of Glucose-Lowering Pharmacotherapy Using Belgian Health Expenditure Data}''. \\ 
        Submitted to Journal of Machine Learning Research.
\item Claesen, M.$^\star$, Gillard, P.$^\star$, De Smet, F., Callens, M., De Moor, B. \& Mathieu, C. (2015). ``\emph{Mortality in individuals treated with glucose lowering agents: a large, controlled cohort study}''. \\ 
        Submitted to Journal of Clinical Endocrinology and Metabolism.
\item Claesen, M., Davis, J., De Smet, F., \& De Moor, B. (2015). ``\emph{Assessing Binary Classifiers Using Only Positive and Unlabeled Data}'', \\
        Submitted to Journal of Machine Learning Research.
\item Claesen, M., Simm, J., Popovic, D., Moreau, Y., \& De Moor, B. (2015). ``\emph{Easy hyperparameter search using Optunity}''. \\
        Submitted to Journal of Machine Learning Research.
\end{itemize}

\section*{International conference papers (published)}
\begin{itemize}
\item Claesen, M., \& De Moor, B. (2015). ``\emph{Hyperparameter Search in Machine Learning}''. In Proceedings of the 11th Metaheuristics International Conference (MIC), Agadir, Morocco. \\ Manuscript available at \url{http://arxiv.org/abs/1502.02127}.
\end{itemize}

\section*{Conference abstracts}
\begin{itemize}
\item Claesen, M. \& De Moor, B. (2013). ``\emph{An overview of the EnsembleSVM software package}''. In International Workshop on Technical Computing for Machine Learning and Mathematical Engineering. Leuven, Belgium.
\item Claesen, M, Simm, J., Popovic, D. (2013). \& De Moor, B. ``\emph{Hyperparameter tuning in Python using Optunity}''. In International Workshop on Technical Computing for Machine Learning and Mathematical Engineering. 2014. Leuven, Belgium.
\item Claesen, M. \& De Moor, B. (2014). ``\emph{Efficient Ensemble Learning With Support Vector Machines}'' BENELEARN. Brussels, Belgium.
\end{itemize}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Keep the following % otherwise \includeonly includes empty pages.

% vim: tw=70 nocindent expandtab foldmethod=marker foldmarker={{{}{,}{}}}
